---
title: "GLM with Train/Val/Test"
author: "Bryan Chang"
date: "2025-12-05"
output: html_document
---
# Load packages
```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(pROC)
```
# Plotting fn
```{r}
#' Plots the gcamp response to a stimulation
#' 
#' Plots the time series of a neuron's gcamp before and after a stimulation.
#' 
#' @param worm_ind The worm index (character)
#' @param vol The volume number of the stimulation
#' @param neuron The neuron label whose response we want to plot
#' @param sec_before The number of seconds before the stimulation to plot
#' @param sec_after The number of seconds after the stimulation to plot
#' @param use_lm Logical. If TRUE, fits an interrupted time series model. If FALSE, plots means.
#' @param sec_before_fit The number of seconds before the stimulation to use for fitting (defaults to sec_before)
#' @param sec_after_fit The number of seconds after the stimulation to use for fitting (defaults to sec_after)
#' 
#' @return A ggplot object
plot_stim_response <- function(
    worm_ind, 
    vol, 
    neuron, 
    sec_before = 30, 
    sec_after = 30, 
    use_lm = FALSE, 
    sec_before_fit = NULL, 
    sec_after_fit = NULL
) {
  # Set default fitting windows
  if (is.null(sec_before_fit)) sec_before_fit <- sec_before
  if (is.null(sec_after_fit)) sec_after_fit <- sec_after
  
  # Desired window of volume indices
  desired_vols <- seq(vol - 2 * sec_before, vol + 2 * sec_after)
  
  # Valid index range for this worm
  max_vol <- nrow(gcamps[[worm_ind]])
  
  # Clip to valid indices
  valid_vols <- desired_vols[desired_vols >= 1 & desired_vols <= max_vol]
  
  if (length(valid_vols) == 0) {
    stop(str_glue("No valid volumes for worm {worm_ind}, vol {vol}."))
  }
  
  # Full time axis (one per desired volume)
  times_full <- seq(-sec_before, sec_after, length.out = length(desired_vols))
  
  # Time axis corresponding only to valid volumes
  times <- times_full[desired_vols %in% valid_vols]
  
  # Extract gcamp
  gc <- gcamps[[worm_ind]][valid_vols, neuron]
  
  df <- tibble(times = times, gc = gc)
  
  # Define before/after sets safely
  before_idx <- which(df$times < 0)
  after_idx  <- which(df$times > 0)
  
  # Fit model if requested
  if (use_lm) {
    df_fit <- df %>% 
      filter(times >= -sec_before_fit & times <= sec_after_fit)
    
    if (nrow(df_fit) < 3) {
      warning("Not enough points to fit lm; skipping.")
    } else {
      fit <- lm(gc ~ times + pmax(times, 0) + I(times > 0), data = df_fit)
      df$fitted <- predict(fit, newdata = df)
    }
  }
  
  # Base plot
  p <- ggplot(df, aes(x = times, y = gc)) +
    geom_line() +
    geom_point() +
    geom_vline(xintercept = 0) +
    labs(x = "Time since stimulation (s)",
         y = "raw gcamp",
         title = str_glue("Worm {worm_ind}, neuron {neuron} response to vol {vol}"))
  
  if (use_lm && "fitted" %in% names(df)) {
    # Add fitted line
    p <- p + geom_line(aes(y = fitted), color = "blue", linewidth = 1)
    
  } else {
    # Mean segments (only if both sides have data)
    if (length(before_idx) > 1) {
      p <- p + geom_segment(
        aes(x    = min(df$times[before_idx]),
            xend = max(df$times[before_idx]),
            y    = mean(df$gc[before_idx]),
            yend = mean(df$gc[before_idx]))
      )
    }
    if (length(after_idx) > 1) {
      p <- p + geom_segment(
        aes(x    = min(df$times[after_idx]),
            xend = max(df$times[after_idx]),
            y    = mean(df$gc[after_idx]),
            yend = mean(df$gc[after_idx]))
      )
    }
  }
  
  p
}
```


# Load data
Download the folder raw_extracted_data from https://osf.io/e2syt/files
```text
ll/
├── explore-sp-atlas
│   └── glm.Rmd
└── raw_extracted_data
```
```{r load data}
# generate worm indices vector
worm_indices <- fs::dir_ls(path = "../raw_extracted_data") %>% # get files in raw_extracted_data
  str_subset("stim_neurons.txt") %>% # keep files including "stim_neurons.txt"
  str_extract("(\\d+)") %>% # extract worm indices
  as.numeric() %>% # convert worm indices from character to numeric
  sort() # sort worm indices

# ensure worm indices are 0, 1, 2, ...
stopifnot(all(worm_indices == seq(0, length(worm_indices) - 1)))

# load stimulation events
stims <- worm_indices %>% 
  map(~ {
    stim_neuron_file <- str_glue("../raw_extracted_data/{.x}_stim_neurons.txt")
    labels_file <- str_glue("../raw_extracted_data/{.x}_labels.txt")
    vols_file <- str_glue("../raw_extracted_data/{.x}_stim_volume_i.txt")
    labels <- readLines(labels_file)
    dat <- tibble(
      vol = as.numeric(readLines(vols_file)) + 1,
      neuron_indices = as.numeric(readLines(stim_neuron_file))
    ) %>% 
      mutate(
        neuron_indices = if_else(neuron_indices < 0,
                                 neuron_indices,
                                 neuron_indices + 1) # increment neuron indices to use 1 indexing
      )
    dat$neuron <- rep(NA, nrow(dat))
    ii <- dat$neuron_indices
    dat$neuron[ii > 0] <- labels[ii[ii > 0]] # give names to stimulated neurons
    return(dat)
  })

names(stims) <- worm_indices
# merge list of tibbles into a single tibble
stim_events <- stims %>% bind_rows(.id = "worm")
```
```{r load gcamp recordings, message=FALSE}
# load gcamp recordings
gcamps <- worm_indices %>% map(~ {
  gcamp <- read_delim(str_glue("../raw_extracted_data/{.x}_gcamp.txt"),
                      delim = " ", col_names = FALSE)
  times <- readr::read_csv(str_glue("../raw_extracted_data/{.x}_t.txt"),
                           col_names = FALSE)[[1]]
  labels <- readLines(str_glue("../raw_extracted_data/{.x}_labels.txt"))
  ii_nolabels <- which(labels %in% c("", " "))
  labels[ii_nolabels] <- str_glue("w{.x}_{ii_nolabels}")
  
  if (ncol(gcamp) != length(labels)) {
    cat(str_glue("Worm {.x} has {length(labels)} labels but {ncol(gcamp)} gcamp columns."), fill = TRUE)
    return(NA)
  }
  
  gcamp <- as.matrix(gcamp)
  rownames(gcamp) <- times
  colnames(gcamp) <- labels
  gcamp
})

names(gcamps) <- worm_indices

# average repeated neurons and remove merge/target neurons
gcamps <- worm_indices %>%
  as.character() %>%
  map(~ {
    ii_keep <- which(!str_detect(colnames(gcamps[[.x]]), "^merge|^target"))
    gcamps[[.x]] <- gcamps[[.x]][, ii_keep]
    repeated <- gcamps[[.x]] %>%
      colnames() %>%
      table() %>%
      as_tibble() %>%
      set_names(c("neuron", "n")) %>%
      filter(n > 1) %>%
      pull(neuron)
    
    for (neur in repeated) {
      ii <- which(colnames(gcamps[[.x]]) == neur)
      gcamps[[.x]][, ii[1]] <- rowMeans(gcamps[[.x]][, ii], na.rm = TRUE)
      gcamps[[.x]] <- gcamps[[.x]][, -ii[-1]]
    }
    return(gcamps[[.x]])
  }) %>% 
  set_names(worm_indices)
```
```{r set aside worms for testing and validation}
set.seed(42)
# proportion of worms
p_train <- 0.8 * 0.8
p_val   <- 0.8 * 0.2
p_test  <- 0.2

# shuffle worms
worms_shuffled <- sample(worm_indices)

n <- length(worms_shuffled)
n_train <- floor(p_train * n)
n_val   <- floor(p_val   * n)
n_test  <- n - n_train - n_val

train_worms <- worms_shuffled[1:n_train]
val_worms   <- worms_shuffled[(n_train+1):(n_train+n_val)]
test_worms  <- worms_shuffled[(n_train+n_val+1):n]

train_idx <- worm_indices %in% train_worms
val_idx   <- worm_indices %in% val_worms
test_idx  <- worm_indices %in% test_worms
```
# Wilcoxon Rank-Sum Test
```{r wilcoxon_test}
TIME_BEFORE <- 20 # 20 vols is equivalent to 10 seconds
TIME_AFTER <- 60

stim_events <- stim_events %>%
  mutate(wilcox_pvals = map2(worm, vol, ~ { # for each row, get worm and vol
    worm_ind <- .x
    vol <- .y
    vols_before <- seq(vol - TIME_BEFORE, vol - 1) # calculate window before stim
    vols_after <- seq(vol + 1, vol + TIME_AFTER) # calculate window after stim
    
    # skip edge cases
    if (all(is.na(gcamps[[worm_ind]]))) {
      cat(str_glue("Worm {worm_ind} is NA."), fill = TRUE)
      return(NA)
    }
    
    if (min(vols_before) < 1 || max(vols_after) > nrow(gcamps[[worm_ind]])) {
      cat(str_glue("Skip: Stim too close to recording boundary."), fill = TRUE)
      return(NA)
    }
    
    gc_before <- gcamps[[worm_ind]][vols_before, , drop = FALSE]
    gc_after <- gcamps[[worm_ind]][vols_after, , drop = FALSE]
    
    pvals <- sapply(seq_len(ncol(gc_before)), function(i) {
      x_before <- gc_before[, i]
      x_after <- gc_after[, i]
      
      tryCatch({
        test <- wilcox.test(x_before, x_after, alternative = "less")
        test$p.value
      }, error = function(e) NA)
    })
    
    list(pvals)
  }))
```
# Acquire features (mean and slope before and after stim) for logistic regression response detector
```{r calculate_features}
TIME_BEFORE_F <-60
TIME_AFTER_F <- 60
TIME_BEFORE_SLOPE <- 20
TIME_AFTER_SLOPE <- 20

stim_events <- stim_events %>%
  rowwise() %>%
  mutate(
    F0 = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      neuron_names <- colnames(worm_data)
      
      if (stim_vol <= 1) return(list(setNames(numeric(), character())))
      
      vols_before <- seq_len(stim_vol - 1)
      if (stim_vol > TIME_BEFORE_F) vols_before <- (stim_vol - TIME_BEFORE_F):(stim_vol - 1)
      
      f0_vals <- sapply(neuron_names, function(nrn) {
        mean(worm_data[vols_before, nrn], na.rm = TRUE)
      })
      
      list(f0_vals)
    },
    
    F = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      neuron_names <- colnames(worm_data)
      total_vols <- nrow(worm_data)
      
      vols_after <- (stim_vol + 1):(stim_vol + TIME_AFTER_F)
      vols_after <- vols_after[vols_after <= total_vols]
      
      if (length(vols_after) == 0) { # can't get mean of 0 points
        return(list(setNames(rep(NA_real_, length(neuron_names)), neuron_names)))
      }
      
      f_vals <- sapply(neuron_names, function(nrn) {
        mean(worm_data[vols_after, nrn], na.rm = TRUE)
      })
      
      list(f_vals)
    },
    
    dFF = {
      f0_vec <- F0
      f_vec <- F
      dff_vals <- (f_vec - f0_vec) / f0_vec
      list(dff_vals)
    },
    
    slope_before = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      neuron_names <- colnames(worm_data)
      times <- as.numeric(rownames(worm_data))
      
      vols_before <- (stim_vol - TIME_BEFORE_SLOPE):(stim_vol - 1)
      vols_before <- vols_before[vols_before > 0]
      
      if (length(vols_before) < 2) {
        return(list(setNames(rep(NA_real_, length(neuron_names)), neuron_names)))
      }
      
      slope_vals <- sapply(neuron_names, function(nrn) {
        y <- worm_data[vols_before, nrn]
        x <- times[vols_before]
        valid <- !is.na(y)
        if (sum(valid) < 2) return(NA_real_) # can't get slope of <2 points
        
        fit <- lm(y[valid] ~ x[valid])
        coef(fit)[2]
      })
      
      list(slope_vals)
    },
    
    slope_after = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      neuron_names <- colnames(worm_data)
      times <- as.numeric(rownames(worm_data))
      total_vols <- nrow(worm_data)
      
      vols_after <- (stim_vol + 1):(stim_vol + TIME_AFTER_SLOPE)
      vols_after <- vols_after[vols_after <= total_vols]
      
      if (length(vols_after) < 2) {
        return(list(setNames(rep(NA_real_, length(neuron_names)), neuron_names)))
      }
      
      slope_vals <- sapply(neuron_names, function(nrn) {
        y <- worm_data[vols_after, nrn]
        x <- times[vols_after]
        valid <- !is.na(y)
        if (sum(valid) < 2) return(NA_real_)
        
        fit <- lm(y[valid] ~ x[valid])
        coef(fit)[2]
      })
      
      list(slope_vals)
    }
  ) %>%
  ungroup()
```

Let's make a feature "rank" that records the number of stimulations have occurred prior to the current one for this worm:
```{r rank feature}
stim_events <- stim_events %>% 
  group_by(worm) %>% 
  mutate(rank = rank(vol) - 1) %>% 
  ungroup()
```
And "stim_rank" that records number of stimulations of itself prior to the current one for this worm:
```{r stim_rank feature}
stim_events <- stim_events %>%
  group_by(worm, neuron) %>%
  arrange(vol) %>%
  mutate(stim_rank = row_number() - 1) %>%
  ungroup()
```
# Boolean indicates whether there are gaps
```{r}
TIME_BEFORE <- 60
TIME_AFTER <- 60
MAX_CONSECUTIVE_NAS <- 2
stim_events <- stim_events %>%
  mutate(has_no_gaps = map2(worm, vol, ~ {
    worm_ind <- .x
    vol      <- .y

    vols_before <- seq(vol - TIME_BEFORE, vol - 1)
    vols_after  <- seq(vol + 1, vol + TIME_AFTER)

    if (all(is.na(gcamps[[worm_ind]]))) return(NA)
    if (min(vols_before) < 1 || max(vols_after) > nrow(gcamps[[worm_ind]])) return(NA)

    gc_before <- gcamps[[worm_ind]][vols_before, , drop = FALSE]
    gc_after  <- gcamps[[worm_ind]][vols_after,  , drop = FALSE]

    sapply(seq_len(ncol(gc_before)), function(i) {
      x_before <- gc_before[, i]
      x_after  <- gc_after[,  i]

      rle_before <- rle(is.na(x_before))
      rle_after  <- rle(is.na(x_after))

      ok_before <- !any(rle_before$values & rle_before$lengths >= MAX_CONSECUTIVE_NAS)
      ok_after  <- !any(rle_after$values  & rle_after$lengths  >= MAX_CONSECUTIVE_NAS)

      ok_before && ok_after
    })
  }))
```
# Expand stim events to long format
```{r expand_to_long}
stim_events_long <- stim_events %>%
  mutate(event_id = row_number()) %>%
  transmute(
    event_id,
    worm,
    vol,
    stim_neuron = neuron,
    rank,
    stim_rank,
    neuron_indices,
    data = pmap(list(F0, F, dFF, slope_before, slope_after, 
                     wilcox_pvals, has_no_gaps), 
                function(f0, f, dff, sb, sa, wp, g) {
      tibble(
        recorded_neuron = names(f0),
        F0 = unname(f0),
        F = unname(f),
        dFF = unname(dff),
        slope_before = unname(sb),
        slope_after = unname(sa),
        wilcox_pval = if(!is.null(wp) && length(wp) > 0) unname(wp[[1]]) else NA,
        has_no_gaps = unname(g)
      )
    })
  ) %>%
  unnest(data)
```

```{r add train/test/val splits}
stim_events_long <- stim_events_long %>%
  mutate(split = case_when(
    worm %in% train_worms ~ "train",
    worm %in% val_worms   ~ "val",
    worm %in% test_worms  ~ "test",
    TRUE                  ~ NA_character_
  ))
```

# Train Binomial Logistic Regression
```text
ll/
├── explore-sp-atlas
│   └── glm.Rmd
└── rdata/gt-table.Rdata
```
```{r train response detector}
load("../rdata/gt-table.Rdata")  # ground-truth labels

# Merge stim_events_long with ground-truth labels
training_data <- stim_events_long %>%
  left_join(
    gt_table %>% select(worm, vol, recorded_neuron, label),
    by = c("worm", "vol", "recorded_neuron")
  ) %>%
  filter(
    !is.na(dFF), !is.na(slope_before), !is.na(slope_after),
    is.finite(dFF), is.finite(slope_before), is.finite(slope_after)
  )

# Keep only 0/1 labels
# -1 labels are negative responses
labeled_data <- training_data %>%
  filter(label %in% c(0, 1)) %>%
  mutate(label = factor(label, levels = c(0, 1)))

# Split: train = model fitting, val = test data, test = use later
train_set <- labeled_data %>% filter(split == "train")
val_set   <- labeled_data %>% filter(split == "val")
# test_set <- labeled_data %>% filter(split == "test")   # will use later

# marker so we know which rows were used for fitting
train_set_marker <- train_set %>%
  select(worm, vol, recorded_neuron) %>%
  mutate(in_train = TRUE)

# Fit logistic regression on training data
binary_model <- glm(
  label ~ dFF + slope_before + slope_after,
  data   = train_set,
  family = binomial()
)

# Predict probabilities for all rows
training_data <- training_data %>%
  mutate(
    logreg_prob = predict(binary_model, newdata = ., type = "response"),
    logreg_pred = ifelse(logreg_prob > 0.5, 1, 0)
  )

# Compute quantile thresholds
q_lo <- quantile(training_data$logreg_prob, 0.05, na.rm = TRUE)
q_hi <- quantile(training_data$logreg_prob, 0.95, na.rm = TRUE)

training_data <- training_data %>%
  mutate(
    logreg_prob = predict(binary_model, newdata = ., type = "response"),
    logreg_pred = ifelse(logreg_prob > 0.5, 1, 0),

    logreg_extreme = case_when(
      logreg_prob < q_lo ~ 0,
      logreg_prob > q_hi ~ 1,
      TRUE ~ NA_real_
    )
  )


# Mark rows that were part of labeled training data
training_data <- training_data %>%
  left_join(train_set_marker,
            by = c("worm", "vol", "recorded_neuron")) %>%
  mutate(in_train = replace_na(in_train, FALSE))

cat(str_glue("Total rows in training_data: {nrow(training_data)}\n\n"))
n_in_train <- sum(training_data$in_train)
cat(str_glue("Number of samples used for training: {n_in_train}\n"))
```

```{r Are stimulated neurons likely to be recorded as responding?}
training_data %>%
  group_by(split) %>%        # split = "train", "val", "test"
  summarize(
    frac_same = mean(logreg_pred[stim_neuron == recorded_neuron], na.rm = TRUE),
    frac_diff = mean(logreg_pred[stim_neuron != recorded_neuron], na.rm = TRUE)
  ) %>%
  filter(split %in% c("train", "val"))
```

```{r plot those with extreme p-vals}
# Sample a few highs and lows
extreme_1 <- training_data %>% 
  filter(logreg_extreme == 1) %>% 
  slice_head(n = 5)

extreme_0 <- training_data %>% 
  filter(logreg_extreme == 0) %>% 
  slice_head(n = 5)

# Plot function wrapper for convenience
plot_example <- function(row) {
  plot_stim_response(
    worm_ind = row$worm,
    vol      = row$vol,
    neuron   = row$recorded_neuron,   # adjust if you want stim_neuron instead
    sec_before = 20,
    sec_after  = 20,
    use_lm = TRUE                     # or FALSE if you prefer the mean-segment version
  )
}

# Generate lists of plots
plots_extreme_1 <- purrr::map(split(extreme_1, seq_len(nrow(extreme_1))), plot_example)
plots_extreme_0 <- purrr::map(split(extreme_0, seq_len(nrow(extreme_0))), plot_example)
```

```{r noticed that the p-vals close to 1 tend to be negative responses}
plots_extreme_1[[4]]
plots_extreme_0[[5]]
```



# Get connectome features
```{r get connectome graphs}
# 1. Load base anatomical connectome
connectome <- read_delim("../connectome/aconnectome_witvliet_2020_8.csv")

# 2. Chemical connectome (directed)
chem_edges <- connectome %>% 
  filter(type == "chemical") %>%
  select(pre, post, synapses)

g_chem <- graph_from_data_frame(
  chem_edges,
  directed = TRUE
)

dist_chem <- distances(
  g_chem,
  mode = "out",
  weights = NA   # unweighted shortest paths
)

# 3. Electrical connectome (undirected)
elec_edges <- connectome %>% 
  filter(type == "electrical") %>%
  select(pre, post, synapses)

g_elec <- graph_from_data_frame(
  elec_edges,
  directed = FALSE
)

dist_elec <- distances(
  g_elec,
  mode = "out",
  weights = NA
)

# 4. Neuropeptidergic connectome
np_df <- read.csv(
  "../connectome/01022024_neuropeptide_connectome_mid_range_model.csv",
  row.names = 1,
  check.names = FALSE
)

np_mat <- as.matrix(np_df)

g_np <- graph_from_adjacency_matrix(
  np_mat,
  mode = "directed",
  weighted = NULL,
  diag = FALSE
)

dist_np <- distances(
  g_np,
  mode = "out",
  weights = NA
)

# 5. Safe distance utilities (separate per connectome)
safe_dist_single <- function(mat) {
  # Returns a function stim, rec -> distance
  function(stim, rec) {
    if (!(stim %in% rownames(mat)) || !(rec %in% colnames(mat))) {
      return(NA_real_)
    }
    d <- mat[stim, rec]
    if (!is.finite(d)) return(NA_real_)
    d
  }
}

# Build graph-specific safe distance functions
safe_dist_chem  <- safe_dist_single(dist_chem)
safe_dist_elec  <- safe_dist_single(dist_elec)
safe_dist_np    <- safe_dist_single(dist_np)

# Vectorized versions for dplyr
safe_dist_vec_chem <- Vectorize(safe_dist_chem)
safe_dist_vec_elec <- Vectorize(safe_dist_elec)
safe_dist_vec_np   <- Vectorize(safe_dist_np)
```
# Predict the binary response using a GLM
```{r create df for glm model with binary classes}
# Compute response of each stim neuron to itself
self_response <- training_data %>%
  filter(stim_neuron == recorded_neuron) %>%
  group_by(stim_neuron, worm, vol) %>%
  summarize(
    stim_self_responded = as.integer(any(logreg_pred == 1, na.rm = TRUE)),
    .groups = "drop"
  )

model_df_bin <- training_data %>%
  left_join(self_response,
            by = c("stim_neuron", "worm", "vol")) %>%
  filter(
    !is.na(stim_neuron),
    stim_neuron != "",
    !grepl("^\\s*$", stim_neuron)    # catches whitespace-only
  ) %>%
  mutate(
    # stim_neuron never had a self-recording row for that worm+vol
    stim_self_responded = replace_na(stim_self_responded, 0)
  ) %>%
  mutate(
    wilcox_pred = as.integer(wilcox_pval < 0.05)
  ) %>%
  transmute(
    rank,
    stim_rank,
    logreg_pred = as.factor(logreg_pred),
    logreg_extreme = as.factor(logreg_extreme),
    wilcox_pred = as.factor(wilcox_pred),
    stim_self_responded,
    stim_neuron,
    recorded_neuron,
    worm,
    vol,
    split,
  ) %>%
  filter(!is.na(logreg_pred))

model_df_bin <- model_df_bin %>%
  mutate(
    is_self = as.integer(stim_neuron == recorded_neuron),
    
    dist_chem  = safe_dist_vec_chem(stim_neuron, recorded_neuron),
    dist_elec  = safe_dist_vec_elec(stim_neuron, recorded_neuron),
    dist_np    = safe_dist_vec_np(stim_neuron, recorded_neuron),

    inv_dist_chem = if_else(is.finite(dist_chem) & dist_chem > 0,
                            1 / dist_chem,
                            0),

    inv_dist_elec = if_else(is.finite(dist_elec) & dist_elec > 0,
                            1 / dist_elec,
                            0),

    inv_dist_np   = if_else(is.finite(dist_np) & dist_np > 0,
                            1 / dist_np,
                            0)
  )


# Compute the constructed heuristic for residual past stimulations
inv_path_sum_generic <- function(past_neurons, current_neuron, safe_fun) {
  d <- safe_fun(past_neurons, current_neuron)
  sum(ifelse(is.finite(d) & d > 0, 1/d, 0))
}

dt <- as.data.table(model_df_bin)
setorder(dt, worm, vol)

dt[, c("hist_inv_chem_30s",
       "hist_inv_elec_30s",
       "hist_inv_np_30s") :=
      {
        n <- .N
        out_chem <- numeric(n)
        out_elec <- numeric(n)
        out_np   <- numeric(n)

        for (i in seq_len(n)) {
          idx <- which(
            vol < vol[i] &
            (vol[i] - vol) < 80 &
            logreg_pred == 1
          )

          if (length(idx) > 0) {
            past <- recorded_neuron[idx]
            cur  <- recorded_neuron[i]

            out_chem[i] <- inv_path_sum_generic(past, cur, safe_dist_vec_chem)
            out_elec[i] <- inv_path_sum_generic(past, cur, safe_dist_vec_elec)
            out_np[i]   <- inv_path_sum_generic(past, cur, safe_dist_vec_np)
          }
        }

        list(out_chem, out_elec, out_np)
      },
      by = worm
]

model_df_bin <- as.data.frame(dt)
```

```{r add reversed chemical conn}
dist_chem_rev <- distances(g_chem, mode = "in", weights = NA)
safe_dist_chem_rev <- safe_dist_single(dist_chem_rev)
safe_dist_vec_chem_rev <- Vectorize(safe_dist_chem_rev)
model_df_bin <- model_df_bin %>%
  mutate(
    dist_chem_rev = safe_dist_vec_chem_rev(recorded_neuron, stim_neuron),

    inv_dist_chem_rev = if_else(
      is.finite(dist_chem_rev) & dist_chem_rev > 0,
      1 / dist_chem_rev,
      0
    )
  )
```
```{r ...and history}
dt[, c("hist_inv_chem_30s",
       "hist_inv_elec_30s",
       "hist_inv_np_30s",
       "hist_inv_chem_rev_30s") :=
      {
        n <- .N
        out_chem     <- numeric(n)
        out_elec     <- numeric(n)
        out_np       <- numeric(n)
        out_chem_rev <- numeric(n)

        for (i in seq_len(n)) {
          idx <- which(
            vol < vol[i] &
            (vol[i] - vol) < 80 &
            logreg_pred == 1
          )

          if (length(idx) > 0) {
            past <- recorded_neuron[idx]
            cur  <- recorded_neuron[i]

            out_chem[i]     <- inv_path_sum_generic(past, cur, safe_dist_vec_chem)
            out_elec[i]     <- inv_path_sum_generic(past, cur, safe_dist_vec_elec)
            out_np[i]       <- inv_path_sum_generic(past, cur, safe_dist_vec_np)

            # reversed chemical direction
            out_chem_rev[i] <- inv_path_sum_generic(past, cur, safe_dist_vec_chem_rev)
          }
        }

        list(out_chem, out_elec, out_np, out_chem_rev)
      },
      by = worm
]
model_df_bin <- as.data.frame(dt)
```


```{r wilcox, rank only}
logreg_model <- glm(
  wilcox_pred ~ rank + is_self,
  data = model_df_bin %>% filter(split == "train"),
  family = binomial(link = "logit")
)

model_df_bin$pred_prob <- predict(logreg_model, newdata = model_df_bin, type = "response")

library(pROC)

# Train ROC
train_df <- model_df_bin %>% filter(split == "train")
roc_train <- roc(train_df$logreg_pred, train_df$pred_prob)
auc_train <- auc(roc_train)

# Val ROC
val_df <- model_df_bin %>% filter(split == "val")
roc_val <- roc(val_df$logreg_pred, val_df$pred_prob)
auc_val <- auc(roc_val)

plot(roc_train, col = "blue",  lwd = 2, main = "ROC: Train vs Val")
lines(roc_val, col = "red", lwd = 2)
legend("bottomright",
       legend = c(paste0("Train AUC = ", round(auc_train, 3)),
                  paste0("Val AUC = ", round(auc_val, 3))),
       col = c("blue", "red"),
       lwd = 2)

# View model summary
summary(logreg_model)
```
```{r logreg, rank only}
logreg_model <- glm(
  logreg_pred ~ rank + is_self,
  data = model_df_bin %>% filter(split == "train"),
  family = binomial(link = "logit")
)

model_df_bin$pred_prob <- predict(logreg_model, newdata = model_df_bin, type = "response")

library(pROC)

# Train ROC
train_df <- model_df_bin %>% filter(split == "train")
roc_train <- roc(train_df$logreg_pred, train_df$pred_prob)
auc_train <- auc(roc_train)

# Val ROC
val_df <- model_df_bin %>% filter(split == "val")
roc_val <- roc(val_df$logreg_pred, val_df$pred_prob)
auc_val <- auc(roc_val)

plot(roc_train, col = "blue",  lwd = 2, main = "ROC: Train vs Val")
lines(roc_val, col = "red", lwd = 2)
legend("bottomright",
       legend = c(paste0("Train AUC = ", round(auc_train, 3)),
                  paste0("Val AUC = ", round(auc_val, 3))),
       col = c("blue", "red"),
       lwd = 2)

# View model summary
summary(logreg_model)
```

```{r marginal coefficients}
features <- c("rank", "stim_rank", "is_self", "inv_dist_chem", "inv_dist_elec", "inv_dist_np",
              "inv_dist_chem_rev",
              "hist_inv_chem_30s", "hist_inv_elec_30s", "hist_inv_np_30s", "stim_self_responded",
              "hist_inv_chem_rev_30s")

marginal_auc <- function(df, feature) {
  
  # Special handling for stim_self_responded to exclude predicting the feature
  if (feature == "stim_self_responded") {
    df2 <- df %>% filter(stim_neuron != recorded_neuron)
    if (nrow(df2) == 0) return(NA_real_)
    return(auc(roc(df2$logreg_pred, df2[[feature]])))
  }
  
  # Default case
  auc(roc(df$logreg_pred, df[[feature]]))
}

train_df <- model_df_bin %>% filter(split == "train")
val_df   <- model_df_bin %>% filter(split == "val")

results <- tibble(
  feature = features,
  auc_train = map_dbl(features, ~ marginal_auc(train_df, .x)),
  auc_val   = map_dbl(features, ~ marginal_auc(val_df,   .x))
)

print(results)

marginal_auc_extreme <- function(df, feature) {
  
  # Exclude missing extreme labels
  df_clean <- df %>% filter(!is.na(logreg_extreme))
  if (nrow(df_clean) == 0) return(NA_real_)
  
  # Special-case for stim_self_responded: exclude stim = recorded
  if (feature == "stim_self_responded") {
    df2 <- df_clean %>% filter(stim_neuron != recorded_neuron)
    if (nrow(df2) == 0) return(NA_real_)
    return(auc(roc(df2$logreg_extreme, df2[[feature]])))
  }
  
  # Default
  auc(roc(df_clean$logreg_extreme, df_clean[[feature]]))
}

train_df_ext <- model_df_bin %>% filter(split == "train")
val_df_ext   <- model_df_bin %>% filter(split == "val")

results_extreme <- tibble(
  feature   = features,
  auc_train = map_dbl(features, ~ marginal_auc_extreme(train_df_ext, .x)),
  auc_val   = map_dbl(features, ~ marginal_auc_extreme(val_df_ext,   .x))
)

print(results_extreme)
```

