---
title: "ROC and Second Classifier"
author: "Bryan Chang"
date: "2025-10-20"
output: html_document
---
```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(nnet)
library(caret)
library(pROC)
```

# Load Data
```{r load data}
worm_indices <- fs::dir_ls(path = "../raw_extracted_data") %>%
  str_subset("stim_neurons.txt") %>% 
  str_extract("(\\d+)_stim", group = 1) %>% 
  as.numeric() %>% 
  sort()

stopifnot(all(worm_indices == seq(0, length(worm_indices) - 1)))

stims <- worm_indices %>% 
  map(~ {
    stim_neuron_file <- str_glue("../raw_extracted_data/{.x}_stim_neurons.txt")
    labels_file <- str_glue("../raw_extracted_data/{.x}_labels.txt")
    vols_file <- str_glue("../raw_extracted_data/{.x}_stim_volume_i.txt")
    labels <- readLines(labels_file)
    dat <- tibble(
      vol = as.numeric(readLines(vols_file)) + 1,
      neuron_indices = as.numeric(readLines(stim_neuron_file))
    ) %>% 
      mutate(
        neuron_indices = if_else(neuron_indices < 0,
                                 neuron_indices,
                                 neuron_indices + 1)
      )
    dat$neuron <- rep(NA, nrow(dat))
    ii <- dat$neuron_indices
    dat$neuron[ii > 0] <- labels[ii[ii > 0]]
    return(dat)
  })

names(stims) <- worm_indices
stim_events <- stims %>% bind_rows(.id = "worm")
```

```{r load gcamps, message=FALSE}
gcamps <- worm_indices %>% map(~ {
  gcamp <- read_delim(str_glue("../raw_extracted_data/{.x}_gcamp.txt"),
                      delim = " ", col_names = FALSE)
  times <- readr::read_csv(str_glue("../raw_extracted_data/{.x}_t.txt"),
                           col_names = FALSE)[[1]]
  labels <- readLines(str_glue("../raw_extracted_data/{.x}_labels.txt"))
  ii_nolabels <- which(labels %in% c("", " "))
  labels[ii_nolabels] <- str_glue("w{.x}_{ii_nolabels}")
  
  if (ncol(gcamp) != length(labels)) {
    cat(str_glue("Worm {.x} has {length(labels)} labels but {ncol(gcamp)} gcamp columns."), fill = TRUE)
    return(NA)
  }
  
  gcamp <- as.matrix(gcamp)
  rownames(gcamp) <- times
  colnames(gcamp) <- labels
  gcamp
})

names(gcamps) <- worm_indices

# Average repeated neurons and remove merge/target
gcamps <- worm_indices %>%
  as.character() %>%
  map(~ {
    ii_keep <- which(!str_detect(colnames(gcamps[[.x]]), "^merge|^target"))
    gcamps[[.x]] <- gcamps[[.x]][, ii_keep]
    repeated <- gcamps[[.x]] %>%
      colnames() %>% 
      table() %>% 
      as_tibble() %>%
      set_names(c("neuron", "n")) %>%
      filter(n > 1) %>% 
      pull(neuron)
    
    for (neur in repeated) {
      ii <- which(colnames(gcamps[[.x]]) == neur)
      gcamps[[.x]][, ii[1]] <- rowMeans(gcamps[[.x]][, ii], na.rm = TRUE)
      gcamps[[.x]] <- gcamps[[.x]][, -ii[-1]]
    }
    return(gcamps[[.x]])
  }) %>% 
  set_names(worm_indices)
```

# Method 1: Wilcoxon Rank-Sum Test
```{r wilcoxon_test}
TIME_BEFORE <- 20
TIME_AFTER <- 60
MAX_CONSECUTIVE_NAS <- 2

stim_events <- stim_events %>%
  mutate(wilcox_pvals = map2(worm, vol, ~ {
    worm_ind <- .x
    vol <- .y
    vols_before <- seq(vol - TIME_BEFORE, vol - 1)
    vols_after <- seq(vol + 1, vol + TIME_AFTER)
    
    if (all(is.na(gcamps[[worm_ind]]))) {
      cat(str_glue("Worm {worm_ind} is NA."), fill = TRUE)
      return(NA)
    }
    
    if (min(vols_before) < 1 || max(vols_after) > nrow(gcamps[[worm_ind]])) {
      cat(str_glue("Skip: Stim too close to recording boundary."), fill = TRUE)
      return(NA)
    }
    
    # Use drop = FALSE to maintain matrix structure
    gc_before <- gcamps[[worm_ind]][vols_before, , drop = FALSE]
    gc_after <- gcamps[[worm_ind]][vols_after, , drop = FALSE]
    
    # Apply over each neuron (column)
    pvals <- sapply(seq_len(ncol(gc_before)), function(i) {
      x_before <- gc_before[, i]
      x_after <- gc_after[, i]
      
      # Check for excessive NAs
      rle_na_before <- rle(is.na(x_before))
      rle_na_after <- rle(is.na(x_after))
      if (any(rle_na_before$values & rle_na_before$lengths >= MAX_CONSECUTIVE_NAS) ||
          any(rle_na_after$values & rle_na_after$lengths >= MAX_CONSECUTIVE_NAS)) {
        return(NA)
      }
      
      # Remove NAs
      x_before <- x_before[!is.na(x_before)]
      x_after <- x_after[!is.na(x_after)]
      
      if (length(x_before) < 5 || length(x_after) < 5) {
        return(NA)
      }
      
      # Wilcoxon test
      tryCatch({
        test <- wilcox.test(x_after, x_before, alternative = "two.sided")
        test$p.value
      }, error = function(e) NA)
    })
    
    list(pvals)
  }))
```

# Method 2: Linear Model Test (20-20 and 20-60)
```{r lm_test_20_20}
stim_events <- stim_events %>%
  mutate(lm_pvals_20_20 = map2(worm, vol, ~ {
    worm_ind <- .x
    vol <- .y
    vols <- seq(vol - 20, vol + 20)
    
    if (all(is.na(gcamps[[worm_ind]]))) {
      return(NA)
    }
    
    if (min(vols) < 1 || max(vols) > nrow(gcamps[[worm_ind]])) {
      return(NA)
    }
    
    gc <- gcamps[[worm_ind]][vols, , drop = FALSE]
    time <- seq(-20, 20)
    
    pvals <- sapply(seq_len(ncol(gc)), function(i) {
      x <- gc[, i]
      
      rle_na <- rle(is.na(x))
      if (any(rle_na$values & rle_na$lengths >= MAX_CONSECUTIVE_NAS)) {
        return(NA)
      }
      
      tryCatch({
        model_null <- lm(x ~ time)
        model_alt <- lm(x ~ time + pmax(time, 0) + I(time > 0))
        
        f_test <- anova(model_null, model_alt)
        f_test[["Pr(>F)"]][2]
      }, error = function(e) NA)
    })
    
    list(pvals)
  }))
```

```{r lm_test_20_60}
stim_events <- stim_events %>%
  mutate(lm_pvals_20_60 = map2(worm, vol, ~ {
    worm_ind <- .x
    vol <- .y
    vols <- seq(vol - 20, vol + 60)
    
    if (all(is.na(gcamps[[worm_ind]]))) {
      return(NA)
    }
    
    if (min(vols) < 1 || max(vols) > nrow(gcamps[[worm_ind]])) {
      return(NA)
    }
    
    gc <- gcamps[[worm_ind]][vols, , drop = FALSE]
    time <- seq(-20, 60)
    
    pvals <- sapply(seq_len(ncol(gc)), function(i) {
      x <- gc[, i]
      
      rle_na <- rle(is.na(x))
      if (any(rle_na$values & rle_na$lengths >= MAX_CONSECUTIVE_NAS)) {
        return(NA)
      }
      
      tryCatch({
        model_null <- lm(x ~ time)
        model_alt <- lm(x ~ time + pmax(time, 0) + I(time > 0))
        
        f_test <- anova(model_null, model_alt)
        f_test[["Pr(>F)"]][2]
      }, error = function(e) NA)
    })
    
    list(pvals)
  }))
```

# Method 3: Logistic Regression Features
```{r calculate_features}
stim_events <- stim_events %>%
  rowwise() %>%
  mutate(
    F0 = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      
      if (stim_vol <= 1) return(list(setNames(numeric(), character())))
      vols_before <- seq_len(stim_vol - 1)
      if (stim_vol > 60) vols_before <- (stim_vol - 60):(stim_vol - 1)
      
      neuron_names <- colnames(worm_data)
      f0_vals <- sapply(neuron_names, function(nrn) {
        mean(worm_data[vols_before, nrn], na.rm = TRUE)
      })
      
      list(f0_vals)
    },
    
    F = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      neuron_names <- colnames(worm_data)
      total_vols <- nrow(worm_data)
      
      vols_after <- (stim_vol + 1):(stim_vol + 60)
      vols_after <- vols_after[vols_after <= total_vols]
      
      if (length(vols_after) == 0) {
        return(list(setNames(rep(NA_real_, length(neuron_names)), neuron_names)))
      }
      
      f_vals <- sapply(neuron_names, function(nrn) {
        mean(worm_data[vols_after, nrn], na.rm = TRUE)
      })
      
      list(f_vals)
    },
    
    dFF = {
      f0_vec <- F0
      f_vec <- F
      dff_vals <- (f_vec - f0_vec) / f0_vec
      list(dff_vals)
    },
    
    slope_before = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      neuron_names <- colnames(worm_data)
      times <- as.numeric(rownames(worm_data))
      
      vols_before <- (stim_vol - 20):(stim_vol - 1)
      vols_before <- vols_before[vols_before > 0]
      
      if (length(vols_before) < 2) {
        return(list(setNames(rep(NA_real_, length(neuron_names)), neuron_names)))
      }
      
      slope_vals <- sapply(neuron_names, function(nrn) {
        y <- worm_data[vols_before, nrn]
        x <- times[vols_before]
        valid <- !is.na(y)
        if (sum(valid) < 2) return(NA_real_)
        
        fit <- lm(y[valid] ~ x[valid])
        coef(fit)[2]
      })
      
      list(slope_vals)
    },
    
    slope_after = {
      worm_data <- gcamps[[as.character(worm)]]
      stim_vol <- vol
      neuron_names <- colnames(worm_data)
      times <- as.numeric(rownames(worm_data))
      total_vols <- nrow(worm_data)
      
      vols_after <- (stim_vol + 1):(stim_vol + 20)
      vols_after <- vols_after[vols_after <= total_vols]
      
      if (length(vols_after) < 2) {
        return(list(setNames(rep(NA_real_, length(neuron_names)), neuron_names)))
      }
      
      slope_vals <- sapply(neuron_names, function(nrn) {
        y <- worm_data[vols_after, nrn]
        x <- times[vols_after]
        valid <- !is.na(y)
        if (sum(valid) < 2) return(NA_real_)
        
        fit <- lm(y[valid] ~ x[valid])
        coef(fit)[2]
      })
      
      list(slope_vals)
    }
  ) %>%
  ungroup()
```

Let's make a feature "rank" that records the number of stimulations have occurred prior to the current one for this worm:
```{r}
stim_events <- stim_events %>% 
  group_by(worm) %>% 
  mutate(rank = rank(vol) - 1) %>% 
  ungroup()
```

```{r}
stim_events <- stim_events %>%
  group_by(worm, neuron) %>%
  arrange(vol) %>%
  mutate(stim_count_for_stim_neuron = row_number() - 1) %>%
  ungroup()
# Note: neuron may be NA for blank stim; handle accordingly if needed
```

# Get connectome features
```{r}
library(igraph)
connectome <- read.delim("../connectome/aconnectome_witvliet_2020_8.csv")

chem_edges <- connectome %>% 
  filter(type == "chemical") %>%
  select(pre, post, synapses)

g_chem <- graph_from_data_frame(
  chem_edges,
  directed = TRUE
)

elec_edges <- connectome %>% 
  filter(type == "electrical") %>%
  select(pre, post, synapses)

elec <- elec_edges
elec_rev <- data.frame(pre = elec$post, post = elec$pre, synapses = elec$synapses)
elec_edges_bidirectional <- rbind(elec, elec_rev)
g_elec <- graph_from_data_frame(elec_edges_bidirectional, directed = TRUE)

g_all <- igraph::union(g_chem, g_elec)

# Directed shortest-path distances
dist_mat <- igraph::distances(
  g_all,
  mode = "out",      # from pre → post direction
  weights = NA       # treat all edges as unweighted; set weights=edge weights if desired
)

# Convert to a data.frame-friendly form
dist_df <- as.data.frame(dist_mat)
dist_df$from <- rownames(dist_df)

```

```{r expand_to_long}
stim_events_long <- stim_events %>%
  mutate(event_id = row_number()) %>%
  transmute(
    event_id,
    worm,
    vol,
    stim_neuron = neuron,
    rank,
    stim_count_for_stim_neuron,
    neuron_indices,
    data = pmap(list(F0, F, dFF, slope_before, slope_after, 
                     wilcox_pvals, lm_pvals_20_20, lm_pvals_20_60), 
                function(f0, f, dff, sb, sa, wp, lm20, lm60) {
      tibble(
        recorded_neuron = names(f0),
        F0 = unname(f0),
        F = unname(f),
        dFF = unname(dff),
        slope_before = unname(sb),
        slope_after = unname(sa),
        wilcox_pval = if(!is.null(wp) && length(wp) > 0) unname(wp[[1]]) else NA,
        lm_pval_20_20 = if(!is.null(lm20) && length(lm20) > 0) unname(lm20[[1]]) else NA,
        lm_pval_20_60 = if(!is.null(lm60) && length(lm60) > 0) unname(lm60[[1]]) else NA
      )
    })
  ) %>%
  unnest(data)
```

# Add connectome features
```{r}
safe_dist <- function(stim, rec, dist_mat) {
  # If either neuron missing → return NA
  if (!(stim %in% rownames(dist_mat)) || !(rec %in% colnames(dist_mat))) {
    return(NA_real_)
  }
  
  d <- dist_mat[stim, rec]
  if (!is.finite(d)) return(NA_real_)  # unreachable
  return(d)
}

safe_dist_vec <- Vectorize(function(a, b) safe_dist(a, b, dist_mat))

stim_events_long <- stim_events_long %>%
  mutate(
    connectome_path_length =
      safe_dist_vec(stim_neuron, recorded_neuron),
    has_path = !is.na(connectome_path_length)
  )
```

# 1. Train a Response Detector to Output Whether or Not a Trace Represents a Response
## Train a Multinomial Logistic Regression Model
```{r load_labels}
load("../rdata/gt-table.Rdata")  # Contains ground truth labels

training_data <- stim_events_long %>%
  left_join(gt_table %>% select(worm, vol, recorded_neuron, label), 
            by = c("worm", "vol", "recorded_neuron")) %>%
  filter(!is.na(dFF), !is.na(slope_before), !is.na(slope_after),
         is.finite(dFF), is.finite(slope_before), is.finite(slope_after))

labeled_data <- training_data %>%
  filter(!is.na(label)) %>%
  mutate(label = factor(label, levels = c(-1, 0, 1)))

# Train/test split
set.seed(457)
train_idx <- createDataPartition(labeled_data$label, p = 0.8, list = FALSE)
train_set <- labeled_data[train_idx, ]
test_set <- labeled_data[-train_idx, ]

# Create marker for training set rows
train_set_marker <- train_set %>%
  select(worm, vol, recorded_neuron) %>%
  mutate(in_train = TRUE)

# Fit multinomial logistic regression
multinom_model <- multinom(label ~ dFF + slope_before + slope_after,
                          data = train_set)

# Get predictions for all data
training_data <- training_data %>%
  filter(!is.na(dFF), !is.na(slope_before), !is.na(slope_after)) %>%
  mutate(
    logreg_pred = predict(multinom_model, newdata = .),
    logreg_prob = predict(multinom_model, newdata = ., type = "probs")
  )

# Get predictions and probability matrix for all data
logreg_pred <- predict(multinom_model, newdata = training_data)
logreg_prob <- predict(multinom_model, newdata = training_data, type = "probs")

# Ensure column names are character labels
colnames(logreg_prob) <- as.character(colnames(logreg_prob))

# Combine with training_data
training_data <- training_data %>%
  mutate(
    logreg_pred = logreg_pred,
    logreg_prob_response = logreg_prob[, "-1"] + logreg_prob[, "1"]
  )

# Mark which rows are in training set
training_data <- training_data %>%
  left_join(train_set_marker, by = c("worm", "vol", "recorded_neuron")) %>%
  mutate(in_train = replace_na(in_train, FALSE))

cat(str_glue("Total rows in training_data: {nrow(training_data)}\n"))
cat(str_glue("Rows in training set: {sum(training_data$in_train)}\n"))
cat(str_glue("Rows NOT in training set: {sum(!training_data$in_train)}\n"))
```
## Train Binomial Logistic Regression (ignores negative responses)
```{r}
load("../rdata/gt-table.Rdata")  # Contains ground truth labels

# Merge labels + basic cleaning
training_data <- stim_events_long %>%
  left_join(gt_table %>% select(worm, vol, recorded_neuron, label),
            by = c("worm", "vol", "recorded_neuron")) %>%
  filter(
    !is.na(dFF), !is.na(slope_before), !is.na(slope_after),
    is.finite(dFF), is.finite(slope_before), is.finite(slope_after)
  )

# Remove all -1 labels entirely
labeled_data <- training_data %>%
  filter(label %in% c(0, 1)) %>%   # <<< key change
  mutate(label = factor(label, levels = c(0, 1)))

# Train/test split
set.seed(457)
train_idx <- createDataPartition(labeled_data$label, p = 0.8, list = FALSE)
train_set <- labeled_data[train_idx, ]
test_set  <- labeled_data[-train_idx, ]

# Mark training rows
train_set_marker <- train_set %>%
  select(worm, vol, recorded_neuron) %>%
  mutate(in_train = TRUE)

# Fit logistic regression (binomial, not multinomial)
binary_model <- glm(
  label ~ dFF + slope_before + slope_after,
  data = train_set,
  family = binomial()
)

# Predict probabilities for all cleaned training_data
training_data <- training_data %>%
  # keep only rows where logistic regression is valid
  filter(
    !is.na(dFF), !is.na(slope_before), !is.na(slope_after),
    is.finite(dFF), is.finite(slope_before), is.finite(slope_after)
  ) %>%
  mutate(
    # prob of label = 1
    logreg_prob = predict(binary_model, newdata = ., type = "response"),
    # predicted class
    logreg_pred = ifelse(logreg_prob > 0.5, 1, 0)
  )

# Mark training set rows
training_data <- training_data %>%
  left_join(train_set_marker, by = c("worm", "vol", "recorded_neuron")) %>%
  mutate(in_train = replace_na(in_train, FALSE))

cat(str_glue("Total rows in training_data: {nrow(training_data)}\n"))
```

```{r Does stim_neuron==recorded_neuron increase probability of response?}
training_data %>%
  summarize(
    frac_same = mean(logreg_pred[stim_neuron == recorded_neuron], na.rm = TRUE),
    frac_diff = mean(logreg_pred[stim_neuron != recorded_neuron], na.rm = TRUE)
  )
```


# 2. Train a Model to Predict Whether Response Will Occur Given Stimulation History
## First, I try to predict logreg p.
```{r}
# Filter out neurons not in connectome and neurons with too much interpolation
valid_neurons <- colnames(dist_mat)

training_data <- training_data %>%
  mutate(
    stim_neuron = as.character(stim_neuron),
    recorded_neuron = as.character(recorded_neuron)
  ) %>%
  filter(
    # Both neurons must be in the connectome
    stim_neuron %in% valid_neurons,
    recorded_neuron %in% valid_neurons,
    
    # # Wilcoxon p-value must be non-missing
    # !is.na(wilcox_pval)
  )
```

```{r}
# Filter out neurons not in connectome and neurons with too much interpolation
valid_neurons <- colnames(dist_mat)

training_data <- training_data %>%
  mutate(
    stim_neuron = as.character(stim_neuron),
    recorded_neuron = as.character(recorded_neuron)
  ) %>%
  filter(
    # Both neurons must be in the connectome
    stim_neuron %in% valid_neurons,
    recorded_neuron %in% valid_neurons,
    
    # Wilcoxon p-value must be non-missing
    !is.na(wilcox_pval)
  )
```

```{r}
model_df <- training_data %>%
  transmute(
    rank,
    stim_count_for_stim_neuron,
    connectome_path_length,
    has_path = as.numeric(has_path),   # convert logical → numeric if needed
    logreg_prob = as.numeric(logreg_prob)
  ) %>%
  filter(!is.na(logreg_prob))
```

```{r}
svg("../figures/prob_dist.svg", width = 10, height = 10)

ggplot(model_df, aes(x = logreg_prob)) +
  geom_histogram(bins = 50, color = "white", fill = "steelblue") +
  labs(
    title = "Histogram of logreg_prob",
    x = "logreg_prob",
    y = "Count"
  ) +
  theme_minimal()

dev.off()
```


```{r}
# Train/test split
set.seed(123)
idx <- createDataPartition(model_df$logreg_prob, p = 0.8, list = FALSE)

trainX <- model_df[idx, ]
testX  <- model_df[-idx, ]

# Example: logistic regression replaced with simple linear model
prob_model <- lm(plogis(logreg_prob) ~ rank,
                 data = trainX)

summary(prob_model)

pred_prob <- predict(prob_model, newdata = testX)
```

```{r}
# Train/test split
set.seed(123)
idx <- createDataPartition(model_df$logreg_prob, p = 0.8, list = FALSE)

trainX <- model_df[idx, ]
testX  <- model_df[-idx, ]

# Example: logistic regression replaced with simple linear model
prob_model <- glm(logreg_prob ~ rank + stim_count_for_stim_neuron +
                   connectome_path_length + has_path,
                 data = trainX, family = binomial)

summary(prob_model)

pred_prob <- predict(prob_model, newdata = testX)
```

```{r}
# Beta model?
library(betareg)

eps <- 1e-6
prob_beta <- pmin(pmax(model_df$logreg_prob, eps), 1 - eps)

beta_model <- betareg(
  prob_beta ~ rank + stim_count_for_stim_neuron +
               connectome_path_length,
  data = model_df
)

summary(beta_model)
```

```{r}
library(mgcv)
gam_model <- gam(
  prob_beta ~ 
    s(rank, k = 5) +
    s(stim_count_for_stim_neuron, k = 5) +
    s(connectome_path_length, k = 5),
  data = model_df
)
summary(gam_model)

svg("../figures/gam_spline.svg", width = 10, height = 10)
plot(gam_model, pages = 1)
dev.off()
```

## It may be more interpretable to predict a binary response with GLM.
```{r creates df for glm model with bin classes}
# Compute response of each stim neuron to itself
self_response <- training_data %>%
  filter(stim_neuron == recorded_neuron) %>%
  group_by(stim_neuron, worm, vol) %>%
  summarize(
    stim_self_responded = as.integer(any(logreg_pred == 1, na.rm = TRUE)),
    .groups = "drop"
  )

model_df_bin <- training_data %>%
  left_join(self_response,
            by = c("stim_neuron", "worm", "vol")) %>%
  mutate(
    # stim_neuron never had a self-recording row for that worm+vol
    stim_self_responded = replace_na(stim_self_responded, 0)
  ) %>%
  transmute(
    rank,
    stim_count_for_stim_neuron,
    connectome_path_length,
    has_path = as.numeric(has_path),   # Convert logical to numeric if needed
    logreg_pred = as.factor(logreg_pred),
    stim_self_responded,
    stim_neuron,
    recorded_neuron,
    worm,
    vol
  ) %>%
  filter(!is.na(logreg_pred)) %>%
  mutate(
    # Compute inverse path length. Consider exponential decay too
    same_node = ifelse(!is.na(connectome_path_length) &
                         connectome_path_length == 0,
                       1, 0),

    inv_dist = ifelse(!is.na(connectome_path_length) &
                        connectome_path_length > 0,
                      1 / connectome_path_length,
                      0)
  ) %>%
  mutate(
    # Impute connectome_path_length to median
    connectome_path_length = case_when(
      is.na(connectome_path_length) ~ median(connectome_path_length, na.rm = TRUE),  # Impute median
      TRUE ~ connectome_path_length  # Keep existing values if not NA
    )
  )

model_df_bin %>%
  distinct(stim_neuron, worm, vol, stim_self_responded) %>%
  summarize(frac_stim_self_responded = mean(stim_self_responded, na.rm = TRUE))

self_response %>%
  summarize(frac_stim_self_responded = mean(stim_self_responded))
```

```{r glm rank only}
# Strangely, glm on rank still has a positive coefficient
# Fit logistic regression
logreg_model <- glm(
  logreg_pred ~ rank + same_node,
  data = model_df_bin,
  family = binomial(link = "logit")
)

# View model summary
summary(logreg_model)

# Predict probabilities for the positive class
predicted_prob <- predict(logreg_model, type = "response")
# Create the ROC curve
roc_curve <- roc(model_df_bin$logreg_pred, predicted_prob)
# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
# Calculate and display AUC
auc(roc_curve)
```
```{r glm with connectome path length}
# Fit logistic regression
logreg_model <- glm(
  logreg_pred ~ rank +
    stim_count_for_stim_neuron +
    connectome_path_length +
    has_path,
  data = model_df_bin,
  family = binomial(link = "logit")
)

# View model summary
summary(logreg_model)

# Predict probabilities for the positive class
predicted_prob <- predict(logreg_model, type = "response")
# Create the ROC curve
roc_curve <- roc(model_df_bin$logreg_pred, predicted_prob)
# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
# Calculate and display AUC
auc(roc_curve)
```
```{r glm with inverse path length and same_node}
logreg_model <- glm(
  logreg_pred ~ rank +
    stim_count_for_stim_neuron +
    inv_dist +
    same_node +
    has_path,
  data = model_df_bin,
  family = binomial(link = "logit")
)

# View model summary
summary(logreg_model)

# Predict probabilities for the positive class
predicted_prob <- predict(logreg_model, type = "response")
# Create the ROC curve
roc_curve <- roc(model_df_bin$logreg_pred, predicted_prob)
# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
# Calculate and display AUC
auc(roc_curve)
```
```{r glm with whether stim neuron had detected response}
# Fit logistic regression
logreg_model <- glm(
  logreg_pred ~ rank +
    stim_count_for_stim_neuron +
    inv_dist +
    same_node +
    has_path +
    stim_self_responded,
  data = model_df_bin,
  family = binomial(link = "logit")
)

# View model summary
summary(logreg_model)

# Predict probabilities for the positive class
predicted_prob <- predict(logreg_model, type = "response")
# Create the ROC curve
roc_curve <- roc(model_df_bin$logreg_pred, predicted_prob)
# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
# Calculate and display AUC
auc(roc_curve)
```
```{r compute list of neurons that had detectable response in the past s seconds}
# this idea is too slow because it computes rowwise
# model_df_bin <- model_df_bin %>%
#   rowwise() %>%
#   mutate(
#     relation_to_historical_response_60s = {
#       # Identify past rows for the same worm
#       past_rows <- model_df_bin %>%
#         filter(
#           worm == cur_data()$worm,
#           logreg_pred == 1,
#           vol < cur_data()$vol,
#           (cur_data()$vol - vol) < 80
#         )
# 
#       past_neurons <- past_rows$recorded_neuron
#       current_neuron <- cur_data()$recorded_neuron
# 
#       if (length(past_neurons) == 0) {
#         0
#       } else {
#         # Compute path lengths using your safe_dist_vec
#         pls <- safe_dist_vec(past_neurons, current_neuron)
# 
#         # Sum inverse path lengths, ignoring NA and zero
#         sum(ifelse(is.finite(pls) & pls > 0, 1 / pls, 0))
#       }
#     }
#   ) %>%
#   ungroup()

dt <- as.data.table(model_df_bin)
setorder(dt, worm, vol)  # ensure chronological order within worm

# Pre-split path-length calculation into a function that accepts vectors
inv_path_sum <- function(past_neurons, current_neuron) {
  pls <- safe_dist_vec(past_neurons, current_neuron)
  sum(ifelse(is.finite(pls) & pls > 0, 1/pls, 0))
}

dt[, relation_to_historical_response_30s := {
  out <- numeric(.N)
  for (i in seq_len(.N)) {
    # restrict to last 80 half-seconds
    idx <- which(vol < vol[i] & (vol[i] - vol) < 80 & logreg_pred == 1)
    if (length(idx) > 0) {
      out[i] <- inv_path_sum(recorded_neuron[idx], recorded_neuron[i])
    }
  }
  out
}, by = worm]

model_df_bin <- as.data.frame(dt)
```
```{r glm with response history in past 40s (1 stim event before with some leeway)}
# Fit logistic regression
logreg_model <- glm(
  logreg_pred ~ rank +
    stim_count_for_stim_neuron +
    inv_dist +
    same_node +
    has_path +
    stim_self_responded +
    relation_to_historical_response_30s,
  data = model_df_bin,
  family = binomial(link = "logit")
)

# View model summary
summary(logreg_model)

# Predict probabilities for the positive class
predicted_prob <- predict(logreg_model, type = "response")
# Create the ROC curve
roc_curve <- roc(model_df_bin$logreg_pred, predicted_prob)
# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
# Calculate and display AUC
auc(roc_curve)
```
```{r integrating neuropeptide connectome}
# Take this file from https://github.com/LidiaRipollSanchez/Neuropeptide-Connectome.git
neuropeptidergic_connectome <- read.csv("../connectome/01022024_neuropeptide_connectome_mid_range_model.csv",
                                        row.names = 1,
                                        check.names = FALSE)

# Convert dataframe to matrix
np_mat <- as.matrix(neuropeptidergic_connectome)

# Build a directed graph. Weighted edges come from the matrix entries.
g_np <- graph_from_adjacency_matrix(
  np_mat,
  mode = "directed",
  weighted = TRUE,
  diag = FALSE
)

# Directed shortest-path distances
dist_mat <- igraph::distances(
  g_all,
  mode = "in",      # from post → pre direction
  weights = NA       # treat all edges as unweighted; set weights=edge weights if desired
)

safe_dist <- function(stim, rec, dist_mat) {
  # If either neuron missing → return NA
  if (!(stim %in% rownames(dist_mat)) || !(rec %in% colnames(dist_mat))) {
    return(NA_real_)
  }
  
  d <- dist_mat[stim, rec]
  if (!is.finite(d)) return(NA_real_)  # unreachable
  return(d)
}

safe_dist_vec <- Vectorize(function(a, b) safe_dist(a, b, dist_mat))

model_df_bin <- model_df_bin %>%
  mutate(
    np_connectome_path_length =
      safe_dist_vec(stim_neuron, recorded_neuron),
    has_np_path = !is.na(np_connectome_path_length)
  )

model_df_bin <- model_df_bin %>%
  mutate(
    np_connectome_path_length =
      safe_dist_vec(stim_neuron, recorded_neuron),

    np_same_node = ifelse(
      !is.na(np_connectome_path_length) & np_connectome_path_length == 0,
      1, 0
    ),

    np_inv_dist = ifelse(
      !is.na(np_connectome_path_length) & np_connectome_path_length > 0,
      1 / np_connectome_path_length,
      0
    )
  )
```
```{r}
# Fit logistic regression
logreg_model <- glm(
  logreg_pred ~ rank +
    stim_count_for_stim_neuron +
    inv_dist +
    same_node +
    has_path +
    stim_self_responded +
    relation_to_historical_response_30s +
    np_inv_dist +
    has_np_path,
  data = model_df_bin,
  family = binomial(link = "logit")
)

# View model summary
summary(logreg_model)

# Predict probabilities for the positive class
predicted_prob <- predict(logreg_model, type = "response")
# Create the ROC curve
roc_curve <- roc(model_df_bin$logreg_pred, predicted_prob)
# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
# Calculate and display AUC
auc(roc_curve)
```
```{r}
# Fit logistic regression
logreg_model <- glm(
  logreg_pred ~ rank +
    stim_count_for_stim_neuron +
    inv_dist +
    same_node +
    has_path +
    stim_self_responded +
    relation_to_historical_response_30s +
    np_inv_dist +
    has_np_path,
  data = model_df_bin,
  family = binomial(link = "logit")
)

# View model summary
summary(logreg_model)

# Predict probabilities for the positive class
predicted_prob <- predict(logreg_model, type = "response")
# Create the ROC curve
roc_curve <- roc(model_df_bin$logreg_pred, predicted_prob)
# Calculate and display AUC
auc(roc_curve)
```



```{r trying to compute SHAP. can do for a single row, but this R method takes too long for full observations}
library(iml)

X <- model_df_bin[, setdiff(names(model_df_bin), "logreg_pred")]
y <- model_df_bin$logreg_pred

predictor <- Predictor$new(
  model = logreg_model,
  data = X,
  y = y,
  predict.fun = function(model, newdata) {
    predict(model, newdata, type = "response")
  }
)

all_sh <- lapply(seq_len(nrow(X)), function(i) {
  Shapley$new(predictor, x = X[i, , drop = FALSE])
})

```


# Plotting Functions

```{r plot_functions}
plot_stim_response <- function(worm_ind, vol, neuron, sec_before = 30, sec_after = 30) {
  tryCatch({
    # Check if neuron exists
    if (!neuron %in% colnames(gcamps[[worm_ind]])) {
      stop(str_glue("Neuron {neuron} not found in worm {worm_ind}"))
    }
    
    total_vols <- nrow(gcamps[[worm_ind]])
    vols <- seq(vol - 2 * sec_before, vol + 2 * sec_after)
    
    # Ensure vols are within valid range
    valid_mask <- vols > 0 & vols <= total_vols
    if (sum(valid_mask) == 0) {
      stop("No valid volumes in range")
    }
    
    vols <- vols[valid_mask]
    
    # Extract gc values
    gc <- gcamps[[worm_ind]][vols, neuron]
    
    # Verify lengths match
    if (length(gc) != length(vols)) {
      stop(str_glue("Length mismatch: gc={length(gc)}, vols={length(vols)}"))
    }
    
    # Create time vector
    time_vals <- seq(-sec_before, sec_after, length.out = length(vols))
    
    # Identify before/after indices
    before <- which(vols < vol)
    after <- which(vols > vol)
    
    # Create base plot
    p <- ggplot(tibble(times = time_vals, gc = gc), aes(x = times, y = gc)) + 
      geom_line() +
      geom_point() +
      geom_vline(xintercept = 0, color = "black", linewidth = 1) +
      labs(x = "Time since stimulation (s)",
           y = "Raw GCaMP",
           title = str_glue("Worm {worm_ind}, neuron {neuron}, vol {vol}")) +
      theme_minimal()
    
    # Add mean lines only if we have data
    if (length(before) > 0) {
      p <- p + geom_segment(aes(x = min(time_vals[before]),
                                xend = max(time_vals[before]),
                                y = mean(gc[before], na.rm = TRUE),
                                yend = mean(gc[before], na.rm = TRUE)),
                           color = "blue", linewidth = 1)
    }
    
    if (length(after) > 0) {
      p <- p + geom_segment(aes(x = min(time_vals[after]),
                                xend = max(time_vals[after]),
                                y = mean(gc[after], na.rm = TRUE),
                                yend = mean(gc[after], na.rm = TRUE)),
                           color = "red", linewidth = 1)
    }
    
    return(p)
  }, error = function(e) {
    # Return empty plot with error message
    ggplot() + 
      annotate("text", x = 0.5, y = 0.5, 
               label = str_glue("ERROR: {e$message}"),
               size = 6, color = "red") +
      labs(title = str_glue("Worm {worm_ind}, neuron {neuron}, vol {vol}")) +
      theme_void()
  })
}

plot_with_lm_fit <- function(worm_ind, vol, neuron, time_before = 20, time_after = 20) {
  tryCatch({
    # Check if neuron exists
    if (!neuron %in% colnames(gcamps[[worm_ind]])) {
      stop(str_glue("Neuron {neuron} not found in worm {worm_ind}"))
    }
    
    total_vols <- nrow(gcamps[[worm_ind]])
    vols <- seq(vol - time_before, vol + time_after)
    
    # Ensure vols are within valid range
    valid_mask <- vols > 0 & vols <= total_vols
    if (sum(valid_mask) < 10) {
      stop("Insufficient valid volumes for fitting")
    }
    
    vols <- vols[valid_mask]
    
    # Extract gc values
    gc <- gcamps[[worm_ind]][vols, neuron]
    
    # Verify lengths match
    if (length(gc) != length(vols)) {
      stop(str_glue("Length mismatch: gc={length(gc)}, vols={length(vols)}"))
    }
    
    # Create time vector relative to stimulation
    time <- vols - vol
    
    # Remove NA values for model fitting
    valid_idx <- !is.na(gc)
    if (sum(valid_idx) < 5) {
      stop("Too many NAs for model fitting")
    }
    
    # Fit models on valid data only
    gc_valid <- gc[valid_idx]
    time_valid <- time[valid_idx]
    
    model_null <- lm(gc_valid ~ time_valid)
    model_alt <- lm(gc_valid ~ time_valid + pmax(time_valid, 0) + I(time_valid > 0))
    
    # Create prediction data frame
    pred_df <- data.frame(time_valid = time)
    pred_null <- predict(model_null, newdata = pred_df)
    pred_alt <- predict(model_alt, newdata = pred_df)
    
    # Verify all vectors have same length
    if (length(time) != length(gc) || length(time) != length(pred_null) || length(time) != length(pred_alt)) {
      stop(str_glue("Length mismatch: time={length(time)}, gc={length(gc)}, pred_null={length(pred_null)}, pred_alt={length(pred_alt)}"))
    }
    
    # Create plot with lines connecting data points
    ggplot(tibble(time = time, gc = gc, pred_null = pred_null, pred_alt = pred_alt), 
           aes(x = time, y = gc)) +
      geom_line(color = "gray50") +  # Line connecting actual data points
      geom_point() +
      geom_line(aes(y = pred_null), color = "blue", linetype = "dashed", linewidth = 1) +
      geom_line(aes(y = pred_alt), color = "red", linewidth = 1) +
      geom_vline(xintercept = 0, linewidth = 1) +
      labs(x = "Time (frames)",
           y = "Raw GCaMP",
           title = str_glue("LM Test: Worm {worm_ind}, neuron {neuron}, vol {vol}"),
           subtitle = "Gray = data, Blue dashed = null model, Red = change-point model") +
      theme_minimal()
  }, error = function(e) {
    # Return empty plot with error message
    ggplot() + 
      annotate("text", x = 0.5, y = 0.5, 
               label = str_glue("ERROR: {e$message}"),
               size = 6, color = "red") +
      labs(title = str_glue("LM Test: Worm {worm_ind}, neuron {neuron}, vol {vol}"),
           subtitle = "Gray = data, Blue dashed = null model, Red = change-point model") +
      theme_void()
  })
}

plot_with_features <- function(worm_ind, vol, neuron, features_row) {
  tryCatch({
    p1 <- plot_stim_response(worm_ind, vol, neuron, sec_before = 20, sec_after = 20)
    
    p1 + labs(subtitle = str_glue(
      "dFF={round(features_row$dFF, 3)}, slope_before={round(features_row$slope_before, 4)}, slope_after={round(features_row$slope_after, 4)}\nPred={features_row$logreg_pred}, P(response)={round(features_row$logreg_prob_response, 3)}"
    ))
  }, error = function(e) {
    # Return empty plot with error message
    ggplot() + 
      annotate("text", x = 0.5, y = 0.5, 
               label = str_glue("ERROR: {e$message}"),
               size = 6, color = "red") +
      labs(title = str_glue("Worm {worm_ind}, neuron {neuron}, vol {vol}"),
           subtitle = str_glue(
             "dFF={round(features_row$dFF, 3)}, slope_before={round(features_row$slope_before, 4)}, slope_after={round(features_row$slope_after, 4)}\nPred={features_row$logreg_pred}, P(response)={round(features_row$logreg_prob_response, 3)}"
           )) +
      theme_void()
  })
}
```

# Sample Random Positives
```{r sample_positives}
set.seed(123)

# Define positive threshold
P_THRESHOLD <- 0.05
N_SAMPLES <- 250

# First, filter to only observations with valid values for ALL tests
# and exclude training set
valid_for_all_tests <- training_data %>%
  filter(
    !in_train,  # Exclude training set
    !is.na(wilcox_pval),
    !is.na(lm_pval_20_20),
    !is.na(lm_pval_20_60),
    !is.na(dFF),
    !is.na(slope_before),
    !is.na(slope_after),
    !is.na(logreg_pred),
    !is.na(logreg_prob_response)
  )

cat(str_glue("Total observations valid for all tests (excluding training): {nrow(valid_for_all_tests)}\n"))

# Method 1: Wilcoxon
wilcox_positives <- valid_for_all_tests %>%
  filter(wilcox_pval < P_THRESHOLD) %>%
  sample_n(min(N_SAMPLES, n()))

cat(str_glue("Wilcoxon positives sampled: {nrow(wilcox_positives)}\n"))

# Method 2: LM 20-20
lm_20_20_positives <- valid_for_all_tests %>%
  filter(lm_pval_20_20 < P_THRESHOLD) %>%
  sample_n(min(N_SAMPLES, n()))

cat(str_glue("LM 20-20 positives sampled: {nrow(lm_20_20_positives)}\n"))

# Method 3: LM 20-60
lm_20_60_positives <- valid_for_all_tests %>%
  filter(lm_pval_20_60 < P_THRESHOLD) %>%
  sample_n(min(N_SAMPLES, n()))

cat(str_glue("LM 20-60 positives sampled: {nrow(lm_20_60_positives)}\n"))

# Method 4: Logistic Regression
logreg_positives <- valid_for_all_tests %>%
  filter(logreg_pred %in% c("-1", "1")) %>%
  sample_n(min(N_SAMPLES, n()))

cat(str_glue("LogReg positives sampled: {nrow(logreg_positives)}\n"))

# Create assessment table
assessment_table <- bind_rows(
  wilcox_positives %>% mutate(method = "Wilcoxon", sample_id = row_number()),
  lm_20_20_positives %>% mutate(method = "LM_20_20", sample_id = row_number()),
  lm_20_60_positives %>% mutate(method = "LM_20_60", sample_id = row_number()),
  logreg_positives %>% mutate(method = "LogReg", sample_id = row_number())
) %>%
  mutate(
    manual_label = NA_character_,  # For you to fill in
    notes = NA_character_
  ) %>%
  select(method, sample_id, worm, vol, stim_neuron, recorded_neuron,
         wilcox_pval, lm_pval_20_20, lm_pval_20_60, 
         logreg_pred, logreg_prob_response,
         dFF, slope_before, slope_after,
         label, manual_label, notes, in_train)

# Verify no training samples and all have complete data
cat(str_glue("\nTraining samples in assessment table: {sum(assessment_table$in_train)}\n"))
cat(str_glue("Rows with any NA test values: {sum(is.na(assessment_table$wilcox_pval) | is.na(assessment_table$lm_pval_20_20) | is.na(assessment_table$lm_pval_20_60))}\n"))

# Save assessment table
write_csv(assessment_table, "assessment_table.csv")
```

# Plot Sampled Positives
```{r plot_wilcoxon, warning=FALSE, fig.width=10, fig.height=6}
for (i in 1:nrow(wilcox_positives)) {
  row <- wilcox_positives[i, ]
  
  tryCatch({
    p <- plot_stim_response(row$worm, row$vol, row$recorded_neuron, sec_before = 20, sec_after = 20) +
      labs(subtitle = str_glue(
        "Method: Wilcoxon | p-value = {round(row$wilcox_pval, 4)}\nStim neuron: {row$stim_neuron} | True label: {ifelse(is.na(row$label), 'Unknown', row$label)}"
      ))
    print(p)
  }, error = function(e) {
    message("⚠️ Skipped: worm=", row$worm, " vol=", row$vol, 
            " neuron=", row$recorded_neuron, " (error: ", e$message, ")")
  })
}
```
```{r plot_lm_20_20, warning=FALSE, fig.width=10, fig.height=6}
for (i in 1:nrow(lm_20_20_positives)) {
  row <- lm_20_20_positives[i, ]
  
  tryCatch({
    p <- plot_with_lm_fit(row$worm, row$vol, row$recorded_neuron, 20, 20) +
      labs(subtitle = str_glue(
        "Method: LM 20-20 | p-value = {round(row$lm_pval_20_20, 4)}\nStim neuron: {row$stim_neuron} | True label: {ifelse(is.na(row$label), 'Unknown', row$label)}"
      ))
    print(p)
  }, error = function(e) {
    message("⚠️ Skipped (LM 20-20): worm=", row$worm, 
            " vol=", row$vol, 
            " neuron=", row$recorded_neuron, 
            " | error: ", e$message)
  })
}
```

```{r plot_lm_20_60, warning=FALSE, fig.width=10, fig.height=6}
for (i in 1:nrow(lm_20_60_positives)) {
  row <- lm_20_60_positives[i, ]
  
  tryCatch({
    p <- plot_with_lm_fit(row$worm, row$vol, row$recorded_neuron, 20, 60) +
      labs(subtitle = str_glue(
        "Method: LM 20-60 | p-value = {round(row$lm_pval_20_60, 4)}\nStim neuron: {row$stim_neuron} | True label: {ifelse(is.na(row$label), 'Unknown', row$label)}"
      ))
    print(p)
  }, error = function(e) {
    message("⚠️ Skipped (LM 20-60): worm=", row$worm, 
            " vol=", row$vol, 
            " neuron=", row$recorded_neuron, 
            " | error: ", e$message)
  })
}
```

```{r plot_logreg, warning=FALSE, fig.width=10, fig.height=6}
for (i in 1:nrow(logreg_positives)) {
  row <- logreg_positives[i, ]
  
  tryCatch({
    p <- plot_with_features(row$worm, row$vol, row$recorded_neuron, row) +
      labs(subtitle = str_glue(
        "Method: LogReg | Prediction: {row$logreg_pred} | P(response) = {round(row$logreg_prob_response, 3)}\nStim neuron: {row$stim_neuron} | True label: {ifelse(is.na(row$label), 'Unknown', row$label)}"
      ))
    print(p)
  }, error = function(e) {
    message("⚠️ Skipped (LogReg): worm=", row$worm, 
            " vol=", row$vol, 
            " neuron=", row$recorded_neuron, 
            " | error: ", e$message)
  })
}
```

# ROC Curves
```{r roc_analysis}
# Get labeled data only
training_data <- read.csv("assessment_table_filled.csv")

labeled_test_data <- training_data %>%
  filter(!is.na(manual_label)) %>%
  mutate(
    is_response = as.numeric(manual_label != 0),  # Binary: 0 = no response, 1 = response
    wilcox_score = 1 - wilcox_pval,  # Convert p-value to score (lower p = higher score)
    lm_20_20_score = 1 - lm_pval_20_20,
    lm_20_60_score = 1 - lm_pval_20_60
  )

# Calculate ROC curves
roc_wilcox <- roc(labeled_test_data$is_response, labeled_test_data$wilcox_score, 
                  na.rm = TRUE)
roc_lm_20_20 <- roc(labeled_test_data$is_response, labeled_test_data$lm_20_20_score,
                    na.rm = TRUE)
roc_lm_20_60 <- roc(labeled_test_data$is_response, labeled_test_data$lm_20_60_score,
                    na.rm = TRUE)
roc_logreg <- roc(labeled_test_data$is_response, labeled_test_data$logreg_prob_response,
                  na.rm = TRUE)

# Plot ROC curves
plot(roc_wilcox, col = "blue", lwd = 2, main = "ROC Curves: Response Detection Methods")
plot(roc_lm_20_20, col = "green", lwd = 2, add = TRUE)
plot(roc_lm_20_60, col = "orange", lwd = 2, add = TRUE)
plot(roc_logreg, col = "red", lwd = 2, add = TRUE)
legend("bottomright", 
       legend = c(
         str_glue("Wilcoxon (AUC = {round(auc(roc_wilcox), 3)})"),
         str_glue("LM 20-20 (AUC = {round(auc(roc_lm_20_20), 3)})"),
         str_glue("LM 20-60 (AUC = {round(auc(roc_lm_20_60), 3)})"),
         str_glue("LogReg (AUC = {round(auc(roc_logreg), 3)})")
       ),
       col = c("blue", "green", "orange", "red"),
       lwd = 2)

# Print AUC values
cat("\nAUC Values:\n")
cat(str_glue("Wilcoxon: {round(auc(roc_wilcox), 3)}\n"))
cat(str_glue("LM 20-20: {round(auc(roc_lm_20_20), 3)}\n"))
cat(str_glue("LM 20-60: {round(auc(roc_lm_20_60), 3)}\n"))
cat(str_glue("LogReg: {round(auc(roc_logreg), 3)}\n"))
```

```{r}
# LR with just dFF
```


```{r save_ROC_plot}
ggroc(list(
  Wilcoxon = roc_wilcox,
  `LM 20-20` = roc_lm_20_20,
  `LM 20-60` = roc_lm_20_60,
  LogReg = roc_logreg
)) +
  theme_minimal() +
  labs(title = "ROC Curves: Response Detection Methods")

# Save as PNG
ggsave("../roc_curves.png", width = 10, height = 6, dpi = 300)
```


```{r save_results}
save(stim_events_long, training_data, assessment_table,
     roc_wilcox, roc_lm_20_20, roc_lm_20_60, roc_logreg,
     file = "../combined_methods_results.Rdata")
```